The framing of active school travel in Ontario, Canada as a health and built environment issue

### 2021-12-01 AST-Framing-Ontario/AST-Framing-Ontario.tex
Discussion and conclusion completed, references updated

\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%%%%%%%%%%%%%%%% end my additions to header

@@

% Pandoc citation processing
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

Rates of walking and bicycling to school, commonly known as active
school travel (AST), have been declining in Canada and the United States
for decades (\textbf{rothmanDeclineActiveSchool2018?}), with levels much
lower than other developed countries like The Netherlands
(\textbf{vangoeverdenSchoolTravelBehaviour2013?}), Denmark
(\textbf{jensenHowObtainHealthy2008?}), and Japan
(\textbf{waygoodChapterSixteenJapan2020?}). However, the appetite for
AST may be high in many Canadian communities. For instance, 40\% of
children in a study conducted in Toronto, Ontario who are driven to
school would like to travel by bicycle instead
(\textbf{laroucheRatherBikeSchool2016a?}). But the study reported that
less than 3\% actually do, despite the vast majority having access to a
bicycle and living within a short and bikeable distance to school.
Another study based in London, Ontario reported similar findings with
respect to children's preference for active travel
(\textbf{larsenRouteBasedAnalysisCapture2012?}). From a public health
perspective, this represents a huge opportunity to improve health and
wellbeing if more children could use active modes to school.

In response to this trend, the Government of Ontario, in Canada, created
a fund in 2017 to support communities across the province in developing
AST initiatives. Green Communities Canada runs the program, called
\emph{Ontario Active School Travel}, and determines the projects that
will be funded with their support (\textbf{GreenCommunities2016?}). As
of December 2021, OAST has funded and provided resources for over 20
projects across Ontario. Many communities implemented school travel
planning (STP), along with encouragement activities like walking school
buses or developed resources for schools and parents, among other
actions (see \textbf{GreenCommunitiesProjects?}). In this context, STP
is a popular ``school-specific'' intervention led by a facilitator that
brings together a committee of stakeholders from diverse sectors
including education, planning, transportation, and public health to
develop action plans (\textbf{buliungSchoolTravelPlanning2011?};
\textbf{mammenSchoolTravelPlanning2014?}). The five-step process
involves identifying barriers to AST based on the local context and
implementing approaches or activities following the 5 E's that alleviate
concerns about AST and make AST safer and more convenient
(\textbf{buliungSchoolTravelPlanning2011?};
\textbf{langUnderstandingModalChoice2011?}).

How AST is framed by STP stakeholders to a target audience, like parents
or the general public, may raise awareness about the issue and influence
how walking and bicycling to school are perceived. Gamson and Modigliani
(\textbf{gamsonChangingCulture1987?}) define a frame as a ``central
organizing idea or story line that provides meaning'' to a particular
phenomenon. A frame can enable individuals ``to locate, perceive,
identify, and label'' information pertaining to various dimensions of an
issue (Goffman, 1974). Therefore, stakeholder groups involved in STP
efforts can play a role in shaping public perception about AST in such a
way that it attracts greater attention and becomes a ``problem'' that
needs to be addressed through behaviour change or by new policies. We
also contend that the ways in which the AST issue is framed can affect
the policy process, as outlined in Kingdon's
(\textbf{kingdonAgendasAlternativesPublic1984?}) Multiple Streams
Framework (MSF), through all three streams (i.e., problem, policy, and
politics). This framing can ultimately impact financial contributions
and potential solutions to address the decline of AST in Canada.

@@

The benefits of AST are valuable information to communicate to the
public in order to convey the importance of this issue. The desire to
increase AST in Canada is certainly warranted - there is compelling
evidence that children who actively commute to school earn physical and
mental health benefits. Faulkner et al. (2009) concluded from their
systematic review that children who travel on foot or by bicycle to
school generally have higher levels of physical activity than their
peers who are driven to school. However, Schoeppe et al. (2015) reported
no association between AST and physical activity. This relationship
could be dose-dependent, meaning that children would have to travel
longer distances to accumulate physical activity. A walking distance of
1000-1600 metres to school has been found to contribute to overall
levels of physical activity for boys
(\textbf{faulknerSchoolTravelChildren2013?}). The daily routine of
travelling to school can be a good opportunity for children to regularly
build physical activity into their schedule
(\textbf{mitraIndependentMobilityMode2013?}). Research has also shown
that using active modes to school contributes to improvements in
cardiovascular fitness
(\textbf{borrestadExperiencesRandomisedControlled2012?}).

More recently, there has been literature exploring the link between
transport and children's wellbeing
(\textbf{waygoodTransportChildrenWellbeing2020?}), with relevant
applications to the study of travel satisfaction (van den Berg et al.,
2020; \textbf{westmanChildrenTravelSchool2017?}). Being driven to school
reduces community interactions for children which may impact their
social wellbeing (\textbf{waygoodChildrenTravelIncidental2015?}),
whereas walking and bicycling give children opportunities to socialize
with friends or siblings
(\textbf{michailChildrenExperiencesTheir2021?}). This is something that
children highly value about travel
(\textbf{zwertsHowChildrenView2010a?}) and indicates that social
connections through travel are important for their wellbeing. AST also
provides opportunities for children to engage with the natural
environment (\textbf{fuscoUnderstandingChildrenPerceptions2012?};
\textbf{romeroChildrenExperiencesEnjoyment2015?}) which can improve
their mental or emotional wellbeing.

@@

Factors that influence AST have been presented and organized using a
socio-ecological model (SEM)
(\textbf{mitraIndependentMobilityMode2013?}) or systems model
(\textbf{badlandDevelopmentSystemsModel2016?}) whereby children's travel
behaviour is understood within the context of their household, social,
neighbourhood, and policy environments. Individual characteristics of
the child are also determinants. The SEM comes from the field of public
health and is a useful framework for understanding complex health
behaviours, including travel habits such as walking or bicycling to
school, because it identifies multiple determinants that need to be
addressed by interventions to facilitate behaviour change. The consensus
in the literature is that various factors are interrelated and that
interventions should target multiple levels in order to increase levels
of healthy and physically active modes of travel to school
(\textbf{mitraIndependentMobilityMode2013?}). Rather than describing an
extensive list of factors that influence AST and mode choice in Canada,
which has been covered elsewhere (e.g., Mammen et al., 2012; Wilson et
al., 2018; \textbf{mitraIndependentMobilityMode2013?};
\textbf{rothmanDeclineActiveSchool2018?}), we briefly discuss a few
potentially modifiable determinants that may be targeted for change
through STP.

At the individual level, older child age is often associated with AST
(Mammen et al., 2012; Wilson et al., 2018;
\textbf{starkExploringIndependentActive2018?}) as parents typically
escort younger children. There is some evidence that gender is a
determinant of AST (\textbf{larsenInfluencePhysicalEnvironment2009?}),
with boys being more likely to travel using active modes than girls,
although this is not a strong or consistent finding (Schoeppe et al.,
2015; \textbf{rothmanDeclineActiveSchool2018?}). Children's mode choice
to school is strongly influenced by their parents' travel behaviours and
the complexity of their household's travel needs (Buliung et al., 2021),
which indicates that shifting parental perceptions and habits is
important. Convenience and inclement weather have been cited by parents
as barriers to AST (\textbf{buliungSchoolTravelPlanning2011?}). Parental
perceptions of the built or school environment (De Meester et al., 2014;
Panter et al., 2010) and their children's skills (Ghekiere et al., 2017)
also influence whether they allow their children to walk or bicycle to
school.

Distance between home and school is most strongly associated with AST
(Ikeda et al., 2018; Mammen et al., 2012; Pont et al., 2009;
\textbf{rothmanDeclineActiveSchool2018?}) with less AST reported among
children who have to travel farther to school. Many studies have also
found that the quality of the built environment along the route to
school and around the school site (Ikeda et al., 2018; Kim and Lee,
2020; \textbf{rothmanActiveSchoolTransportation2021?}) and provision of
active travel infrastructure (Pont et al., 2009;
\textbf{chenPromotingActiveStudent2018a?}) facilitate AST. Canadian
youth report that they feel most safe bicycling on streets in their
neighbourhood or that have low volumes of traffic
(\textbf{TACbikeinfra2020?}). Finally, concerns about traffic and
strangers have been reported by parents who drive their children to
school (Mammen et al., 2012), which highlights that the volume and speed
of cars can be a concern or deterrent for AST.

@@

School travel planning (STP) has been implemented in Canada since at
least the late 2000s. Within the STP process, facilitators generally
establish multi-sector committees who intervene at the participating
school through a range of activities related to the 5E's such as
\emph{education} strategies, \emph{encouragement} through in-person
events or programs, \emph{engineering} improvements to or around the
school site, and \emph{enforcement} of traffic speeds around schools
(Mammen et al., 2014; \textbf{langUnderstandingModalChoice2011?}). The
first large-scale evaluation of STP as an intervention in Canada took
place at twelve schools across the country, including four in Ontario,
using parental surveys to measure changes in travel behaviour and
perceptions (\textbf{buliungSchoolTravelPlanning2011?}). Since then,
there have been other assessments of both the efficacy of STP
(Buttazzoni et al., 2019; Mammen et al., 2014) and the process of
implementing such programs in Ontario (Buttazzoni et al., 2018;
\textbf{mammenPuttingSchoolTravel2015?}). STP facilitators have
recommended that additional time and resources are needed to improve the
efficacy of STP (\textbf{mammenPuttingSchoolTravel2015?}), which
highlights that long-term and sustained efforts driven at the policy
level are required to address declining rates of AST.

The ways in which AST is framed seem particularly important to shift
parental attitudes and perceptions given their reported influence on
children's travel mode to school. STP activities heavily focus on
education or encouragement (Buttazzoni et al., 2018; Mammen et al.,
2014; \textbf{buliungSchoolTravelPlanning2011?}), but parents may not
always be receptive to the goals of STP and may be resistant to
behaviour change (see Buttazzoni et al., 2018). Parents have been found
to express different understandings, language, and perceptions than
planners of the built environment determinants that influence school
travel (Buliung et al., 2021). This is also true when it comes to other
factors like convenience of different modes to school
(\textbf{langUnderstandingModalChoice2011?}). STP stakeholder groups
must pay special attention to parents' understanding of the decline of
AST as a problem, which may affect their receptivity to proposed
solutions.

The ``central organizing idea or story line'' of AST, to apply the
definition from Gamson and Modigliani
(\textbf{gamsonChangingCulture1987?}), could also affect broader support
in the community. Municipal representatives are perceived to be
instrumental but the involvement of other stakeholder groups (e.g.,
busing consortium representatives and local residents) can be lacking
(Buttazzoni et al., 2018). In Ontario, it would be reasonable to say
that AST has become a policy issue on the education and public health
agendas over this time as evidenced by the financial contribution from
the provincial government. The support from a range of municipal
representatives (see Buttazzoni et al., 2018;
\textbf{mammenPuttingSchoolTravel2015?}) demonstrates that the ``policy
stream'' (see \textbf{kingdonAgendasAlternativesPublic1984?}) has been
well engaged. It is unknown to what degree the general public (i.e.,
local residents) has been exposed to messaging about AST, which could
affect their participation in the STP process as desired by STP
facilitators (Buttazzoni et al., 2018).

@@

We assembled a collection of publicly available documents that were
sourced online from the main stakeholder groups involved in STP
initiatives in Ontario: i) school boards (public or Catholic and
English-speaking only); ii) municipal governments; and iii)
transportation consortia. The latter are a unique group of entities
sanctioned by Ontario's Ministry of Education in Ontario 2006. Each
consortium involves a collaboration between regions and school boards;
the consortium's objective is to deliver more efficient and timely
transportation services to schools in each region
(\textbf{ministryeducation2006?}). Non-profit organizations, police
services, and advocacy groups are other stakeholders who often play a
role in supporting AST and/or STP, but this study does not include any
documents from these groups because they are not consistently
participating in all initiatives across Ontario.

@@

In the final step, we removed all blank spaces, punctuation,
capitalization, and numbers. English stop words, which are common words
such as \emph{and} or \emph{the} as identified in a predetermined list
by Lewis et al. (\textbf{R-tm?}) and other frequent terms in the
documents like ``school'' and specific location names, were removed from
the corpora.

@@

Issues that pertain to public health or wellbeing are often presented to
the public through particular frames to influence perceptions or
behaviours. As previously mentioned, a frame is a ``central organizing
idea or story line that provides meaning'' to a public issue or
phenomenon (\textbf{gamsonChangingCulture1987?}). Scholars in the field
of political communications have proposed that communicators, such as
the media or an institution, construct the narrative of a frame for
policy positions or public issues in order to activate or restrict a
particular response in the intended audience (Pan and Kosicki, 1993).
Organized groups of stakeholders can employ similar methods to attract
attention to particular issues. Framing can be used to position existing
solutions as suitable to address particular issues (Mah et al., 2014),
which may prevent the public from being aware of other policy approaches
that challenge the status quo. The way policy issues are framed is
ultimately important to understand because it plays a role in either
altering or preserving the existing social perceptions. This, in turn,
can affect whether issues are put on the agenda of policy makers and can
determine which solutions are proposed to address the problem
(\textbf{kingdonAgendasAlternativesPublic1984?}).

Framing of issues is an important step in developing health policy. An
obvious example over the past decade is the framing of climate change as
a public health issue (e.g., Depoux et al., 2017; Maibach et al., 2010;
Weathers and Kendall, 2016) to increase public engagement and awareness
of the issue. This framing has slowly advanced this issue on public
policy agendas as public attention puts pressure on the policy stream to
adopt frameworks for action. For example, transport planners also use
different frames to guide the extent to which transport policies can be
adapted to address climate change. In a recent paper
(\textbf{reynardGrowthResilienceHow2021?}), framing analysis was applied
to review the representation of issues such as mobility and social
exclusion in municipal policies from four western Canadian cities under
the current circumstances of climate change. The authors found four
primary frames: ``The Growing City,'' ``If You Build It, They Will
Come,'' ``Better City for All,'' and a ``the Resilient City''
(\textbf{reynardGrowthResilienceHow2021?}). Each frame presented the
nature, opportunities, and challenges of climate change in different
ways which set the stage for the types of mitigation and adaptation
strategies that cities were proposing to address this issue.

@@

We use topic modelling in R to conduct the framing analysis. Topic
modelling is a machine learning technique that can identify what
language and concepts are being communicated by analyzing text. This
method is more practical for researchers working with large amounts of
text because it replaces the manual coding of topics that would normally
take place to analyze or summarize textual data
(\textbf{jacobiQuantitativeAnalysisLarge2016?}). In the data
pre-processing phase, we tokenize the text in the documents and create a
document-term matrix so that it is in the correct format for analysis.
We primarily use the following packages: \texttt{tidytext}
(\textbf{R-tidytext?}), \texttt{topicmodels} (\textbf{R-topicmodels?}),
\texttt{word2vec} (\textbf{R-word2vec?}), and \texttt{wordcloud}
(\textbf{R-wordcloud?}) to examine text in the documents that were
sourced for this project. These packages have functions for determining
the frequency of specific words in each document or relationships (e.g.,
pairs of adjacent terms called bigrams) and correlations between words.
These methods can reveal what language and concepts are being
communicated to the general public. Topic modelling is a popular method
for analyzing text from social media platforms
(\textbf{albalawiUsingTopicModeling2020?}) or news articles
(\textbf{jacobiQuantitativeAnalysisLarge2016?}). We estimate latent
Dirichlet allocation (LDA) models to classify both the STP and documents
according to the topics that are contained within them. This method
``treats each document as a mixture of topics, and each topic as a
mixture of words'' (\textbf{silgeTextMining2021?}). The model's output
is ``a set of topics consisting of clusters of words that co-occur in
these documents according to certain patterns''
(\textbf{jacobiQuantitativeAnalysisLarge2016?}). Researchers must then
interpret the identified topics, as done after other methods of manual
coding. We also compare the topics between the policy documents and the
academic papers, and use our interpretation to answer the questions
outlined above that guide this paper. This methodology is not familiar
to many, so we describe in more detail below how we used specific
functions and code in R to produce our results.

@@

This paper is an example of open and reproducible research that uses
only open software. All data were obtained from publicly available
sources and organized in the form of a data package. Following best
practices in spatial data science (\textbf{brunsdon2020opening?}), the
code and data needed to reproduce or conduct a similar analysis for
other regions in North America or elsewhere are available for download.

@@

\caption{\label{tab:content-table}\label{tab:policy-concordance}CAPTION THIS TABLE}
\centering
\begin{tabular}[t]{>{}ll>{\raggedright\arraybackslash}p{20em}}
\toprule
Terms & Stakeholder & Context\\
\midrule
\textbf{\cellcolor{gray!6}{Air Quality}} & \cellcolor{gray!6}{School Board} & \cellcolor{gray!6}{Active transportation [...] improves air quality.}\\
\textbf{Benefit} & Municipality & Stronger bones and muscles, improved self-esteem and sense of well-being while reducing stress and risk of chronic disease all benefit those who use active transportation.\\
\textbf{\cellcolor{gray!6}{Walking School Bus}} & \cellcolor{gray!6}{School Board} & \cellcolor{gray!6}{While taking part in a walking school bus, your child will enjoy seeing friends on the way to school. They will be active more often. This is also a great opportunity for your child to socialize with school friends in a monitored and safe way where they can practice social distancing, modelled by a leader.}\\
\textbf{Community} & School Board & Help your students get started on the right foot - encourage them to walk or bike to school when possible. Even leaving the car a block or two and walking the rest of the way helps. It’s good for the environment and your health, and teaches your child independence and community awareness.\\
\textbf{\cellcolor{gray!6}{Emissions}} & \cellcolor{gray!6}{Consortia} & \cellcolor{gray!6}{An active school commute also reduces congestion in school zones and contributes to reducing greenhouse gas emissions – it’s a win-win for everyone!}\\
\addlinespace
\textbf{Health} & Municipality & Active School Travel allows school-aged children the chance to participate in moderate to intense physical activity. This is linked with lower body mass index and improved cardiovascular health.\\
\textbf{\cellcolor{gray!6}{Lanes}} & \cellcolor{gray!6}{Municipality} & \cellcolor{gray!6}{We are continuing to build on the cycling and pedestrian network by adding more bike lanes, building multi-use paths and encouraging developments to provide better pedestrian/cycling environments.}\\
\textbf{Mental Health} & Municipality & ASST not only improves physical and mental health but contributes to a healthier environment and safer streets.\\
\textbf{\cellcolor{gray!6}{Physical Health}} & \cellcolor{gray!6}{Municipality} & \cellcolor{gray!6}{Encouraging Active Transportation promotes personal health and recreation, helps manage congestion, reduces emissions and supports municipal objectives for efficient land use.}\\
\bottomrule
\end{tabular}
\end{table}

>>>>

\usepackage{graphicx}
%%%%%%%%%%%%%%%% end my additions to header

@@

% Pandoc citation processing

@@

Rates of walking and bicycling to school, commonly known as active
school travel (AST), have been declining in Canada and the United States
for decades (Rothman et al., 2018), with levels much lower than other
developed countries like The Netherlands (van Goeverden and de Boer,
2013), Denmark (Jensen, 2008), and Japan (Waygood and Taniguchi, 2020).
However, the appetite for AST may be high in many Canadian communities.
For instance, 40\% of children in a study conducted in Toronto, Ontario
who are driven to school would like to travel by bicycle instead
(Larouche et al., 2016). But the study reported that less than 3\%
actually do, despite the vast majority having access to a bicycle and
living within a short and bikeable distance to school. Another study
based in London, Ontario reported similar findings with respect to
children's preference for active travel (Larsen et al., 2012). From a
public health perspective, this represents a huge opportunity to improve
health and wellbeing if more children could use active modes to school.

In response to this trend, the Government of Ontario, in Canada, created
a fund in 2017 to support communities across the province in developing
AST initiatives. Green Communities Canada runs the program, called
\emph{Ontario Active School Travel}, and determines the projects that
will be funded with their support (Canada, 2020a). As of December 2021,
OAST has awarded over 2 million Canadian dollars and provided resources
for over 25 projects across Ontario. Many communities implemented school
travel planning (STP), along with encouragement activities like walking
school buses or developed resources for schools and parents, among other
actions (see Canada, 2020b). In this context, STP is a popular
``school-specific'' intervention led by a facilitator that brings
together a committee of stakeholders from diverse sectors including
education, planning, transportation, and public health to develop action
plans (Buliung et al., 2011; Mammen, M. R. Stone, Ron Buliung, et al.,
2014). The five-step process involves identifying barriers to AST based
on the local context and implementing approaches or activities following
the 5 E's that alleviate concerns about AST and make AST safer and more
convenient (Buliung et al., 2011; Lang et al., 2011).

How AST is framed by STP stakeholders to a target audience, like parents
or the general public, may raise awareness about the issue and influence
how walking and bicycling to school are perceived. Gamson and Modigliani
(1987) define a frame as a ``central organizing idea or story line that
provides meaning'' to a particular phenomenon. A frame can enable
individuals ``to locate, perceive, identify, and label'' information
pertaining to various dimensions of an issue (Goffman, 1974). Therefore,
stakeholder groups involved in STP efforts can play a role in shaping
public perception about AST in such a way that it attracts greater
attention and becomes a ``problem'' that needs to be addressed through
behaviour change or by new policies. We also contend that the ways in
which the AST issue is framed can affect the policy process, as outlined
in Kingdon's (1984) Multiple Streams Framework (MSF), through all three
streams (i.e., problem, policy, and politics). This framing can
ultimately impact financial contributions and potential solutions to
address the decline of AST in Canada.

@@

The benefits of AST are valuable information to communicate to the
public in order to convey the importance of this issue. The desire to
increase AST in Canada is certainly warranted - there is compelling
evidence that children who actively commute to school earn physical and
mental health benefits. Faulkner et al.~(2009) concluded from their
systematic review that children who travel on foot or by bicycle to
school generally have higher levels of physical activity than their
peers who are driven to school. However, Schoeppe et al.~(2015) reported
no association between AST and physical activity. This relationship
could be dose-dependent, meaning that children would have to travel
longer distances to accumulate physical activity. A walking distance of
1000-1600 metres to school has been found to contribute to overall
levels of physical activity for boys (Faulkner et al., 2013). The daily
routine of travelling to school can be a good opportunity for children
to regularly build physical activity into their schedule (Mitra, 2013).
Research has also shown that using active modes to school contributes to
improvements in cardiovascular fitness (Børrestad et al., 2012).

More recently, there has been literature exploring the link between
transport and children's wellbeing (Waygood et al., 2020), with relevant
applications to the study of travel satisfaction (van den Berg et al.,
2020; Westman, Olsson, et al., 2017). Being driven to school reduces
community interactions for children which may impact their social
wellbeing (Waygood and Friman, 2015), whereas walking and bicycling give
children opportunities to socialize with friends or siblings (Michail et
al., 2021). This is something that children highly value about travel
(Zwerts et al., 2010) and indicates that social connections through
travel are important for their wellbeing. AST also provides
opportunities for children to engage with the natural environment (Fusco
et al., 2012; Romero, 2015) which can improve their mental or emotional
wellbeing.

@@

Factors that influence AST have been presented and organized using a
socio-ecological model (SEM) (Mitra, 2013) or systems model (Badland et
al., 2016) whereby children's travel behaviour is understood within the
context of their household, social, neighbourhood, and policy
environments. Individual characteristics of the child are also
determinants. The SEM comes from the field of public health and is a
useful framework for understanding complex health behaviours, including
travel habits such as walking or bicycling to school, because it
identifies multiple determinants that need to be addressed by
interventions to facilitate behaviour change. The consensus in the
literature is that various factors are interrelated and that
interventions should target multiple levels in order to increase levels
of healthy and physically active modes of travel to school (Mitra,
2013). Rather than describing an extensive list of factors that
influence AST and mode choice in Canada, which has been covered
elsewhere (e.g., Mammen et al., 2012; Mitra, 2013; Rothman et al., 2018;
Wilson et al., 2018), we briefly discuss a few potentially modifiable
determinants that may be targeted for change through STP.

At the individual level, older child age is often associated with AST
(Mammen et al., 2012; Stark et al., 2018; Wilson et al., 2018) as
parents typically escort younger children. There is evidence that gender
is a determinant of AST, with boys being more likely to travel using
active modes than girls, although this is not a strong or consistent
finding (Rothman et al., 2018; Schoeppe et al., 2015). Children's mode
choice to school is strongly influenced by their parents' travel
behaviours and the complexity of their household's travel needs (Buliung
et al., 2021), which indicates that shifting parental perceptions and
habits is important. Convenience and inclement weather have been cited
by parents as barriers to AST (Buliung et al., 2011). Parental
perceptions of the built or school environment (De Meester et al., 2014;
Panter et al., 2010) and their children's skills (Mammen et al., 2012)
also influence whether they allow their children to walk or bicycle to
school.

Distance between home and school is most strongly associated with AST
(Ikeda et al., 2018; Mammen et al., 2012; Pont et al., 2009; Rothman et
al., 2018) with less AST reported among children who have to travel
farther to school. Many studies have also found that the quality of the
built environment along the route to school and around the school site
(Ikeda et al., 2018; Rothman et al., 2021) and provision of active
travel infrastructure (Chen et al., 2018; Pont et al., 2009) facilitate
AST. Canadian youth report that they feel most safe bicycling on streets
in their neighbourhood or that have low volumes of traffic
({\textbf{???}}). Finally, concerns about traffic and strangers have
been reported by parents who drive their children to school (Mammen et
al., 2012), which highlights that the volume and speed of cars can be a
concern or deterrent for AST.

@@

School travel planning (STP) has been implemented in Canada since at
least the late 2000s. Within the STP process, facilitators generally
establish multi-sector committees who intervene at the participating
school through a range of activities related to the 5E's such as
\emph{education} strategies, \emph{encouragement} through in-person
events or programs, \emph{engineering} improvements to or around the
school site, and \emph{enforcement} of traffic speeds around schools
(Lang et al., 2011; Mammen, M. R. Stone, G. Faulkner, et al., 2014). The
first large-scale evaluation of STP as an intervention in Canada took
place at twelve schools across the country, including four in Ontario,
using parental surveys to measure changes in travel behaviour and
perceptions (Buliung et al., 2011). Since then, there have been other
assessments of both the efficacy of STP (Buttazzoni et al., 2019;
Mammen, M. R. Stone, G. Faulkner, et al., 2014) and the process of
implementing such programs in Ontario (Buttazzoni et al., 2018; Mammen
et al., 2015). STP facilitators have recommended that additional time
and resources are needed to improve the efficacy of STP (Mammen et al.,
2015), which highlights that long-term and sustained efforts driven at
the policy level are required to address declining rates of AST.

The ways in which AST is framed seem particularly important to shift
parental attitudes and perceptions given their reported influence on
children's travel mode to school. STP activities heavily focus on
education or encouragement (Buliung et al., 2011; Buttazzoni et al.,
2018; Mammen, M. R. Stone, G. Faulkner, et al., 2014), but parents may
not always be receptive to the goals of STP and may be resistant to
behaviour change (see Buttazzoni et al., 2018). Parents have been found
to express different understandings, language, and perceptions than
planners of the built environment determinants that influence school
travel (Buliung et al., 2021). This is also true when it comes to other
factors like convenience of different modes to school (Lang et al.,
2011). STP stakeholder groups must pay special attention to parents'
understanding of the decline of AST as a problem, which may affect their
receptivity to proposed solutions.

The ``central organizing idea or story line'' of AST, to apply the
definition from Gamson and Modigliani (1987), could also affect broader
support in the community. Municipal representatives are perceived to be
instrumental but the involvement of other stakeholder groups (e.g.,
busing consortium representatives and local residents) can be lacking
(Buttazzoni et al., 2018). In Ontario, it would be reasonable to say
that AST has become a policy issue on the education and public health
agendas over this time as evidenced by the financial contribution from
the provincial government. The support from a range of municipal
representatives (see Buttazzoni et al., 2018; Mammen et al., 2015)
demonstrates that the ``policy stream'' (see Kingdon, 1984) has been
well engaged. It is unknown to what degree the general public (i.e.,
local residents) has been exposed to messaging about AST, which could
affect their participation in the STP process as desired by STP
facilitators (Buttazzoni et al., 2018).

@@

We assembled a collection of publicly available documents that were
sourced online from the main stakeholder groups involved in STP
initiatives in Ontario: i) school boards (public or Catholic and
English-speaking only); ii) municipal governments; and iii)
transportation consortia. The latter are a unique group of entities
sanctioned by Ontario's Ministry of Education in Ontario 2006. Each
consortium involves a collaboration between regions and school boards;
the consortium's objective is to deliver more efficient and timely
transportation services to schools in each region. Non-profit
organizations, police services, and advocacy groups are other
stakeholders who often play a role in supporting AST and/or STP, but
this study does not include any documents from these groups because they
are not consistently participating in all initiatives across Ontario.

@@

In the final step, we removed all blank spaces, punctuation,
capitalization, and numbers. English stop words, which are common words
such as \emph{and} or \emph{the} as identified in a predetermined list
by Lewis et al.~({\textbf{???}}) and other frequent terms in the
documents like ``school'' and specific location names, were removed from
the corpora.

@@

Issues that pertain to public health or wellbeing are often presented to
the public through particular frames to influence perceptions or
behaviours. As previously mentioned, a frame is a ``central organizing
idea or story line that provides meaning'' to a public issue or
phenomenon (Gamson and Modigliani, 1987). Scholars in the field of
political communications have proposed that communicators, such as the
media or an institution, construct the narrative of a frame for policy
positions or public issues in order to activate or restrict a particular
response in the intended audience (Pan and Kosicki, 1993). Organized
groups of stakeholders can employ similar methods to attract attention
to particular issues. Framing can be used to position existing solutions
as suitable to address particular issues (Mah et al., 2014), which may
prevent the public from being aware of other policy approaches that
challenge the status quo. The way policy issues are framed is ultimately
important to understand because it plays a role in either altering or
preserving the existing social perceptions. This, in turn, can affect
whether issues are put on the agenda of policy makers and can determine
which solutions are proposed to address the problem (Kingdon, 1984).

Framing of issues is an important step in developing health policy. An
obvious example over the past decade is the framing of climate change as
a public health issue (e.g., Depoux et al., 2017; Maibach et al., 2010;
Weathers and Kendall, 2016) to increase public engagement and awareness
of the issue. This framing has slowly advanced this issue on public
policy agendas as public attention puts pressure on the policy stream to
adopt frameworks for action. For example, transport planners also use
different frames to guide the extent to which transport policies can be
adapted to address climate change. In a recent paper (Reynard et al.,
2021), framing analysis was applied to review the representation of
issues such as mobility and social exclusion in municipal policies from
four western Canadian cities under the current circumstances of climate
change. The authors found four primary frames: ``The Growing City'',
``If You Build It, They Will Come'', ``Better City for All'', and a
``the Resilient City'' (Reynard et al., 2021). Each frame presented the
nature, opportunities, and challenges of climate change in different
ways which set the stage for the types of mitigation and adaptation
strategies that cities were proposing to address this issue.

@@

We use topic modelling in R to conduct the framing analysis. Topic
modelling is a machine learning technique that can identify what
language and concepts are being communicated by analyzing text. This
method is more practical for researchers working with large amounts of
text because it replaces the manual coding of topics that would normally
take place to analyze or summarize textual data (Jacobi et al., 2016).
In the data pre-processing phase, we tokenize the text in the documents
and create a document-term matrix so that it is in the correct format
for analysis. We primarily use the following packages: \texttt{tidytext}
({\textbf{???}}), \texttt{topicmodels} ({\textbf{???}}),
\texttt{word2vec} ({\textbf{???}}), and \texttt{wordcloud}
({\textbf{???}}) to examine text in the documents that were sourced for
this project. These packages have functions for determining the
frequency of specific words in each document or relationships (e.g.,
pairs of adjacent terms called bigrams) and correlations between words.
These methods can reveal what language and concepts are being
communicated to the general public. Topic modelling is a popular method
for analyzing text from social media platforms (Albalawi et al., 2020)
or news articles (Jacobi et al., 2016). We estimate latent Dirichlet
allocation (LDA) models to classify both the STP and documents according
to the topics that are contained within them. This method ``treats each
document as a mixture of topics, and each topic as a mixture of words''
(Silge and Robinson, 2021). The model's output is ``a set of topics
consisting of clusters of words that co-occur in these documents
according to certain patterns'' (Jacobi et al., 2016). Researchers must
then interpret the identified topics, as done after other methods of
manual coding. We also compare the topics between the policy documents
and the academic papers, and use our interpretation to answer the
questions outlined above that guide this paper. This methodology is not
familiar to many, so we describe in more detail below how we used
specific functions and code in R to produce our results.

@@

This paper is an example of open and reproducible research that uses
only open software. All data were obtained from publicly available
sources and organized in the form of a data package. Following best
practices in spatial data science ({\textbf{???}}), the code and data
needed to reproduce or conduct a similar analysis for other regions in
North America or elsewhere are available for download.

@@

\caption{\label{tab:content-table}\label{tab:policy-concordance}The context of key terms that were identified as common bigrams.}
\centering
\begin{tabular}[t]{>{}ll>{\raggedright\arraybackslash}p{20em}}
\toprule
Terms & Stakeholder & Context\\
\midrule
\textbf{\cellcolor{gray!6}{Air Quality}} & \cellcolor{gray!6}{School Board} & \cellcolor{gray!6}{Active transportation [...] improves air quality.}\\
\textbf{Benefit} & Municipality & Stronger bones and muscles, improved self-esteem and sense of well-being while reducing stress and risk of chronic disease all benefit those who use active transportation.\\
\textbf{\cellcolor{gray!6}{Walking School Bus}} & \cellcolor{gray!6}{School Board} & \cellcolor{gray!6}{While taking part in a walking school bus, your child will enjoy seeing friends on the way to school. They will be active more often. This is also a great opportunity for your child to socialize with school friends in a monitored and safe way where they can practice social distancing, modelled by a leader.}\\
\textbf{Community} & School Board & Help your students get started on the right foot - encourage them to walk or bike to school when possible. Even leaving the car a block or two and walking the rest of the way helps. It’s good for the environment and your health, and teaches your child independence and community awareness.\\
\textbf{\cellcolor{gray!6}{Emissions}} & \cellcolor{gray!6}{Consortia} & \cellcolor{gray!6}{An active school commute also reduces congestion in school zones and contributes to reducing greenhouse gas emissions – it’s a win-win for everyone!}\\
\addlinespace
\textbf{Health} & Municipality & Active School Travel allows school-aged children the chance to participate in moderate to intense physical activity. This is linked with lower body mass index and improved cardiovascular health.\\
\textbf{\cellcolor{gray!6}{Lanes}} & \cellcolor{gray!6}{Municipality} & \cellcolor{gray!6}{We are continuing to build on the cycling and pedestrian network by adding more bike lanes, building multi-use paths and encouraging developments to provide better pedestrian/cycling environments.}\\
\textbf{Mental Health} & Municipality & ASST not only improves physical and mental health but contributes to a healthier environment and safer streets.\\
\textbf{\cellcolor{gray!6}{Physical Health}} & \cellcolor{gray!6}{Municipality} & \cellcolor{gray!6}{Encouraging Active Transportation promotes personal health and recreation, helps manage congestion, reduces emissions and supports municipal objectives for efficient land use.}\\
\bottomrule
\end{tabular}
\end{table}
